Determining Frequentist Confidence Limits Using a Directed Parameter Space Search

### 2014-08-04 aps4p5/drafts/draft_140804.tex
new draft added

\title
{An Efficient Parameter Space Search as an Alternative to Markov Chain Monte Carlo} 
\author
{Scott F.\ Daniel$^1$, Andrew J.\ Connolly$^1$, and Jeff Schneider$^2$\\
$^1$Department of Astronomy, University of Washington, Seattle, WA\\ 
$^2$Machine Learning Department, Carnegie Mellon University, Pittsburgh, PA
}
\begin{document}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
We consider the problem of inferring constraints on a high-dimensional parameter
space with a computationally expensive likelihood function.  
Markov chain Monte Carlo (MCMC) methods offer significant improvements in
efficiency over grid-based searches and are easy to implement in a wide range of
cases.  However, MCMCs offer 
few guarantees that all of the interesting regions of 
parameter space are explored.
We propose a
machine learning algorithm that improves upon the performance of MCMC
by intelligently targeting likelihood evaluations so as to quickly and
accurately characterize the likelihood surface in both low- and high-likelihood
regions.  
We compare our algorithm to MCMC on toy examples and the 7-year WMAP
cosmic microwave background data release.  
Our algorithm finds comparable parameter
constraints to MCMC comparably quickly and with greater
certainty that all of the interesting regions of parameter space have been
explored.
\end{abstract}

@@

We propose an alternative solution, using methods from machine learning
(to wit, Gaussian processes, though the algorithm admits many possible drivers; see
Section \ref{sec:gp})
to find the desired confidence limit by directly searching the
parameter space of $\vec{\theta}$ for points which lie on the boundary of the
desired confidence limit $\chi^2\le\chi^2_\text{lim}$.  
We refer to this algorithm as Active Parameter Searching (APS).
We present the algorithm in detail in Sections \ref{sec:algorithm} and
\ref{sec:gp}.
As it searches, APS uses
the knowledge it has already learned about $\vec{\theta}\rightarrow\chi^2$ 
to improve its search according to a metric that rewards the
algorithm both for finding points on the confidence limit and for sampling
points from previously unexplored regions of parameter space.  We find that this
behavior makes APS more robust against multi-modal $\chi^2$ functions (see
Section \ref{sec:toy}) without
sacrificing the speed of convergence 
associated with traditional MCMC methods (see Section \ref{sec:wmap}).
We show that, while APS is designed specifically to
yield Frequentist confidence
limits (which we define in Appendix \ref{sec:frequentism}),
it is also capable of roughly approximating
Bayesian credible limits on the space of
$\vec{\theta}$ (see Sections \ref{sec:bayes},\ref{sec:toy} and \ref{sec:wmap}).
This rough approximation may principally be of use in cases where little 
prior information about $\vec{\theta}$ exists and the researcher
requires APS's robustness against multi-modality to be certain that she
has discovered all of the best fits to her data.

@@

For each point $\vec{M}$ in $\{\hat{\Gamma}\}$, APS performs the following:
\begin{itemize}
\item (1B) If fewer than $N_p$ focused searches have been proposed about $\vec{M}$, 
randomly choose a point from a small sphere surrounding $\vec{M}$ in parameter space and 
evaluate $\chi^2$ at that point.  
\\
\item(2B) If the point does not satisfy $\chi^2=\chi^2_\text{lim}$ (which is highly probable), 
use bisection
along the line connecting the proposed point to $\vec{M}$ to find a point which does
satisfy $\chi^2=\chi^2_\text{lim}$, i.e. find a point on that line for which
$\chi^2<\chi^2_\text{lim}$ -- often $\vec{M}$ itself -- and a point along that line
for which $\chi^2>\chi^2_\text{lim}$ and iteratively step half the distance between them until
arriving at the desired $\chi^2_\text{lim}$ \cite{minuit}.  The points discovered in this way \
with $\chi^2=\chi^2_\text{lim}$ are stored in the set $\{\hat{B}(\vec{M})\}$ of 
``boundary points'' about $\vec{M}$ (each $\vec{M}$ in $\{\hat{\Gamma}\}$ will have its own
set $\{\hat{B}(\vec{M})\}$).  
\\
\item(3B) If $N_p$ boundary points have been found for $\vec{M}$, skip steps (1B) and (2B).
Instead, examine the points in $\{\hat{B}(\vec{M})\}$ and identify the points
corresponding to both the maximum and minimum values of each parameter $\theta_i$.  These
points will be referred to as $\{\hat{B}_\text{MaxMin}(\vec{M})\}$.  
There will be $2\times N_p$ points in $\{\hat{B}_\text{MaxMin}(\vec{M})\}$, one for
the maximum and one for the minimum in each of our $N_p$ parameters.  
\\
\item(4B) For each point $\vec{b}$ in $\{\hat{B}_\text{MaxMin}(\vec{M})\}$, 
propose some number $N_v$ of new points (we have chosen
$N_v=20$ for the sake of speed) $\vec{v} = \vec{b}+\vec{\epsilon}$ where $\vec{\epsilon}$ is a
random, small vector in parameter space constrained to be perpendicular to
$\vec{b}-\vec{M}$.  There will now be $2\times N_p \times N_v$ points $\vec{v}$
($N_v$ such points for each of the $2\times N_p$ extremal boundary points in
$\{\hat{B}_\text{MaxMin}(\vec{M})\}$).
\\
\item(5B) At each point $\vec{v}$ above, find the $N_g$ nearest neighbor points as in step (2A) above
and evaluate equation (\ref{eqn:sstat}).  
Find the point $\vec{v}$ which maximizes $S$ and evaluate $\vec{\theta}\rightarrow\chi^2$ at
that point.  If
$\chi^2 \neq \chi^2_\text{lim}$, bisection is again used and a new boundary point is found.
Store the resulting $\chi^2=\chi^2_\text{lim}$ point in $\{\hat{B}(\vec{M})\}$
\end{itemize}

The modification suggested in section \ref{sec:focus} is designed to explore 
$\chi^2=\chi^2_\text{lim}$ surface by starting from its extreme corners and stepping
perpendicularly to its ``radii'' (the vector $\vec{b}-\vec{M}$ would be a radius of
$\chi^2=\chi^2_\text{lim}$ described a sphere), with the hopes of being able
to extend the bounds of $\{\hat{B}_\text{MaxMin}(\vec{M})\}$.  In this way,
our extended APS attempts to plum the full Frequentist confidence limit $\chi^2\le\chi^2_\text{lim}$
as rapidly as possible.  However, we do not want this to come at the expense of the wide-ranging
exploration of steps (1A-3A).  Our version of APS alternates between performing steps (1A-3A) and steps
(1B-5B).  If there are $N_m$ distinct known local minima in $\{\hat{\Gamma}\}$, 
our algoritm will perform
$N_m$ iterations of steps (1A-3A), and then perform steps (1B-5B) once for each
$\vec{M}$ in $\{\hat{\Gamma}\}$.

@@

The final modification to the original APS algorithm which we propose in this
work is to utilize function minimization to identify local minima in $\chi^2$.
One of the principal drawbacks of the algorithm presented in steps (1A-3A) is
that it makes no account of ``near misses.''  If step (3A) finds that $\chi^2$
is lower than average but still greater than $\chi^2_\text{lim}$, the original
APS algorithm does nothing with that information.  We have tried to improve upon
that with our incorporation of bisection in sections \ref{sec:focus} and
\ref{sec:unitSphere} above.  Our final innovation is to allow APS to use points
with 
\begin{eqnarray}
\chi^2_\text{lim}&\le&\chi^2\le\chi^2_\text{threshold}\label{eqn:simplexCriterion}\\
\chi_\text{threshold}&\equiv&\chi^2_\text{min}+G\times(\chi^2_{1/10}-\chi^2_\text{min})
\label{eqn:threshold}
\end{eqnarray}
to seed a simplex function minimizer as developed by Nelder and Mead (1965).
Here, $G$ is a user-set parameter, $\chi^2_\text{min}$ is the minimum discovered
$\chi^2$ value and $\chi^2_{1/10}$ is the 1/10 quantile of the $\chi^2$ values
discovered by steps (1A-3A) (not those discovered by any of the modifications
discussed in sections \ref{sec:focus}, \ref{sec:unitSphere} or
\ref{sec:simplex}).

@@

@@

We make code implementing our extended APS for 
an arbitrary $\vec{\theta}\rightarrow\chi^2$
function available at \verb|https://github.com/uwssg/APS/|.
The code is written in C++.  It is designed for users to use
with any $\vec{theta}\rightarrow\chi^2$ function.  Users need only
to write a sub-class of the \verb|chisquared| class defined in
\verb|chisq.h| and \verb|chisq.cpp| which evaluates their desired
$\vec{\theta}\rightarrow\chi^2$ and pass an instantiation of this
sub-class to an instantiation of the \verb|aps| class defined in
\verb|aps.h| and \verb|aps.cpp|.  It is this \verb|aps| class which
does the work of performing the searches described above.
The user should not need to modify this class (though she is 
certainly welcome to).

Our \verb|aps| class provides the option for several
user-specified parameters.  We list them below along with the C++ 
methods (belonging to the class \verb|aps|) used to set them.  
This code includes several user-specified parameters which we
outline below.

\begin{itemize}
\item The parameter $G$ from equation (\ref{eqn:threshold}) above can be set using the method
\verb|aps.set_grat|
\\
\item $\chi^2_\text{lim}$ can be manually set using the method \verb|aps.set_target|
\\
\item $\Delta \chi^2$ can be set in the constructor for the class \verb|aps|.
\\
\item The number $N_g$ of nearest neighbors used in step 2A can be set in the constructor
for the class \verb|aps|
\\
\item $N_c$ from step 2A can be set using the method \verb|aps.set_samples|
\\
\item The minimum and maximum bounds in parameter space can be set using the 
method \verb|aps.set_max| and \verb|aps.set_min|.  These bounds will constrain the range of
points sampled in step 2A.  They will not contraint searches performed with bisection or the 
simplex minimizer.
\\
\item If there is a characteristic length scale associated with a parameter, this can be set
using the method \verb|aps.set_characteristic_length|.  This length scale will be used to
normalize all distances in parameter space.  If it is not set, it will default to the
difference between the maximum and minimum values set by \verb|aps.set_max| and
\verb|aps.set_min|.  Setting a characteristic length scale will principally
effect what points are considered nearest neighbors for the purposes of step 2A.
If the characteristic length scale of a parameter is large, that parameter will
be down-weighted when calculating the parameter-space distance between two points.
If the characteristic length scale of a parameter is small, that parameter will
be up-weighted in length calculations.
\end{itemize}

@@

\item(2D) Select the $3N_p+1$ nearest neighbor points to $\vec{a}$ from
$\{\hat{A}\}$ (this will obviously include $\vec{a}$ itself, which can be discarded).  Call
this set of nearest neighbors $\{\hat{N}\}$.
\\
\item(3D) For each point $\vec{n}$ in $\{\hat{N}\}$, calculate the vector
$\vec{\delta}$ such that $\delta_i=|n_i-a_i|$ (here the vertical bars denote the absolute value).  
For each component
$\delta_i$ of $\vec{\delta}$, starting with the largest and working towards the smallest:
\\
\begin{itemize}
\item(3Da) If $n_i>a_i$ and $a_{\text{max},i}$ has not been set, set $a_{\text{max},i}=n_i$.
\\
\item(3Db) If $n_i<a_i$ and $a_{\text{min},i}$ has not been set, set $a_{\text{min},i}=n_i$.
\\
\item(3Dc) If both $a_\text{max,i}$ and $a_\text{min,i}$ have already been set, move on to the
next largest $\delta_i$ and repeat at step (3Da).
\end{itemize}
\item(4D) Choose the next neighbor $\vec{n}$ and repeat from step (3D). Continue until all
$a_{\text{max},i}$ and $a_{\text{min},i}$ have been set.  Once all $a_{\text{max},i}$ and
$a_{\text{min},i}$ have been set, move on to step (5D).
\\
\item(5D) $\vec{a}_\text{max}$ adn $\vec{a}_\text{min}$ should now describe an asymmetric,
rectangular hyperbox about $\vec{a}$.  To symmetrize it, step through all $N_p$ dimensions
and reset either $a_{\text{max},i}$ or $a_{\text{min},i}$ so that 
$d_\text{min}=a_i-a_{\text{min},i}$ and
$d_\text{max}=a_{\text{max},i}-a_i$ are equal to the minimum of $\{d_\text{max},d_\text{min}\}$
from the asymmetric hyperbox.  Return to step (1D) and repeat for the next $\vec{a}$.
\end{itemize}

At this point, each $\vec{a}$ in $\{\hat{A}\}$ should have a corresponding symmetric hyperbox
surrounding it described by the corresponding 
$\vec{a}_\text{max}(\vec{a})$ and $\vec{a}_\text{min}(\vec{a})$.  
To integrate the posterior probability density, we will assume that all of the
parameter space points in each hyperbox correspond have $\chi^2=\chi^2(\vec{a})$
($\chi^2)\vec{a})$ being already known because $\vec{a}$ is one of the points for which APS evaluated
$\vec{theta}\rightarrow\chi^2$).
The posterior probability associated with each box is then
\begin{eqnarray}
P(\vec{a})&=&\frac{V(\vec{a})}{P_\text{tot}}\times\exp\left[-\chi^2(\vec{a})/2\right]
\label{eqn:posterior}\\
P_\text{tot}&\equiv&\sum_{\vec{a}} V(\vec{a})\times\exp\left[-\chi^2(\vec{a})/2\right]
\end{eqnarray}
where $V(\vec{a})$ is the hypervolume of with the hyperbox surrounding $\vec{a}$.
To plot the $(1-\alpha)\%$ credible limit, one need only plot the points in hyperspace that
enclose the lowest-$\chi^2$ hyperboxes containing $(1-\alpha)\%$ of the total probability.
We test this algorithm on a function with an exactly known credible limit
in Section \ref{sec:toy} (see Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}) 
and on a real, physical dataset in Section \ref{sec:wmap} (see Figure \ref{fig:contours_bayes}).  
In  Section \ref{sec:toy}, we find that our results agree with the known credible limit.
In Section \ref{sec:wmap}, we find that our results roughly approximate those derived from traditional
Markov Chain Monte Carlo methods.

@@

Step (2A) of \APS requires that a prediction $\mu$ be made regarding the value
of $\chi^2$ at points as yet unsampled by the algorithm.  An uncertainty $\sigma$
must also be assigned to this prediction.  APS writ large is agnostic regarding
the method used to find $\mu$ and $\sigma$.  We choose to use a Gaussian process.

@@

The normalization constant $K$ in equation (\ref{eqn:likelihood}) --known as the
``Kriging parameter'' for the geophysicist who pioneered the overall method -- 
also
must be assumed.  Note that $K$ also multiplies the diagonal elements
$C_{ii}$.  Determining the value of $K$ 
is somewhat problematic because, examining equation
(\ref{mu}), one sees that the factors of $K$ and $K^{-1}$
cancel out of
the prediction $\mu$, so that the assumed value of $K$ has no effect on the accuracy
of the prediction.  
If the opposite had been true, one could heuristically set $K$ to
maximize the accuracy of $\mu$.  
Instead, we set $K$ to the value that maximizes the likelihood
of the input data, i.e. the $N_g$ nearest neighbors in parameter space
used to perform the Gaussian Process inference.

@@

As APS samples more points in parameter space, we expect it to learn more about the
$\vec{\theta}\rightarrow\chi^2$ function and thus be capable of a more accurate Gaussian
Process, so we allow it to periodically reset the hyper-parameters $\vec{h}$.  However,
minimizing $E$ can be a very time-consuming process, so we have APS keep track of the amount of
time elapsed since the algorithm began.  We divide this value by the number of calls made to
the $\vec{\theta}\rightarrow\chi^2$ function and compare this amortized time to the average
amount of time spent on one call to $\vec{\theta}\rightarrow\chi^2$.  APS us only allowed to
reset its hyper-parameters if the difference between these two times is less than 0.1 second
(of course, we do require that APS optimize its hyper parameters at least once, so that
we can have confidence in our Gaussian Process model).

>>>>

\title
{Determining Frequentist Confidence Limits Using a Directed Parameter Space
Search} 
\author
{Scott F.\ Daniel$^1$, Andrew J.\ Connolly$^1$, and Jeff Schneider$^2$\\
$^1$Department of Astronomy, University of Washington, Seattle, WA\\ 
$^2$Machine Learning Department, Carnegie Mellon University, Pittsburgh, PA
}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
We consider the problem of inferring constraints on a 
high-dimensional parameter
space with a computationally expensive likelihood function.
We propose a machine learning algorithm that maps out the 
Frequentist confidence limit on parameter space by intelligently targeting
likelihood evaluations so as to quickly and accurately characterize the likelihood
surface in both low- and high-likelihood regions.
We compare our algorithm to Bayesian credible limits derived by 
the well-tested Markov Chain Monte Carlo (MCMC)
algorithms using both multi-modal toy likelihood functions and the 
7-year WMAP cosmic microwave background likelihood function.  We find that our
algorithm correctly identifies the location, general size, and general shape of
high-likelihood regions in parameter space while being more robust against
multi-modality than MCMC.
\end{abstract}

@@

@@

We propose an alternative solution, using methods from machine learning
(to wit, Gaussian processes, though the 
algorithm admits many possible drivers; see
Section \ref{sec:gp})
to find the desired confidence limit by directly searching the
parameter space of $\vec{\theta}$ for points which lie on the boundary of the
desired confidence limit $\chi^2\le\chi^2_\text{lim}$.  
We refer to this algorithm as Active Parameter Searching (APS).
We present the algorithm in detail in Sections \ref{sec:algorithm} and
\ref{sec:gp}.
As it searches, APS uses
the knowledge it has already learned about $\vec{\theta}\rightarrow\chi^2$ 
to improve its search according to a metric that rewards the
algorithm both for finding points on the confidence limit and for sampling
points from previously unexplored regions of parameter space.  We find that this
behavior makes APS more robust against multi-modal $\chi^2$ functions (see
Section \ref{sec:toy}) without
sacrificing the speed of convergence 
associated with traditional MCMC methods (see Section \ref{sec:wmap}).
We show that, while APS is designed specifically to
yield Frequentist confidence
limits (which we define in Appendix \ref{sec:frequentism}),
it is also capable of roughly approximating
Bayesian credible limits on the space of
$\vec{\theta}$ (see Sections \ref{sec:bayes},\ref{sec:toy} and \ref{sec:wmap}).

{\bf %marking new text
Readers should be aware that, while MCMC can be analytically shown to
asymptotically sample the Bayesian posterior, APS is designed to discover
Frequentist confidence limits purely through exploration, i.e. there is no {\it a
priori} motivation behind the APS algorithm.  Our only demonstration of APS'
efficacy is empirical.  It has done well on the problems we have tested it on,
though we have tried to choose those problems such that they are representative of
problems likely to be encountered by the community.  Readers who require
theoretically-motivated results will likely need to resort to a sampling routine
like MCMC to derive their final constraints.  We hope to show, however, in Section
\ref{sec:toy}, that even in these use cases, APS can provide a valuable service by
identifying the general regions of parameter space over which the Bayesian
posterior needs to be integrated, i.e. by guiding researchers in the process of setting
priors for a sampling algorithm, a problem which has posed significant problems
for MCMC in the case of multi-modal likelihood functions.
} %marking new text

@@

For each point $\vec{M}$ in $\{\hat{\Gamma}\}$, APS performs the following:
\begin{itemize}
\item (1B) If fewer than $N_p$ focused searches have been proposed about $\vec{M}$, 
randomly choose a point from a small sphere surrounding $\vec{M}$ in parameter space and 
evaluate $\chi^2$ at that point.  
\\
\item(2B) If the point does not satisfy $\chi^2=\chi^2_\text{lim}$ (which is highly probable), 
use bisection
along the line connecting the proposed point to $\vec{M}$ to find a point which does
satisfy $\chi^2=\chi^2_\text{lim}$, i.e. find a point on that line for which
$\chi^2<\chi^2_\text{lim}$ -- often $\vec{M}$ itself -- and a point along that line
for which $\chi^2>\chi^2_\text{lim}$ and iteratively step half the distance between them until
arriving at the desired $\chi^2_\text{lim}$ \cite{minuit}.  The points discovered in this way \
with $\chi^2=\chi^2_\text{lim}$ are stored in the set $\{\hat{B}(\vec{M})\}$ of 
``boundary points'' about $\vec{M}$ (each $\vec{M}$ in $\{\hat{\Gamma}\}$ will have its own
set $\{\hat{B}(\vec{M})\}$).  
\\
\item(3B) If $N_p$ boundary points have been found for $\vec{M}$, skip steps (1B) and (2B).
Instead, examine the points in $\{\hat{B}(\vec{M})\}$ and identify the points
corresponding to both the maximum and minimum values of each parameter $\theta_i$.  These
points will be referred to as $\{\hat{B}_\text{MaxMin}(\vec{M})\}$.  
There will be $2\times N_p$ points in $\{\hat{B}_\text{MaxMin}(\vec{M})\}$, one for
the maximum and one for the minimum in each of our $N_p$ parameters.  
\\
\item(4B) For each point $\vec{b}$ in $\{\hat{B}_\text{MaxMin}(\vec{M})\}$, 
propose some number $N_v$ of new points (we have chosen
$N_v=20$ for the sake of speed) $\vec{v} = \vec{b}+\vec{\epsilon}$ where $\vec{\epsilon}$ is a
random, small vector in parameter space constrained to be perpendicular to
$\vec{b}-\vec{M}$.  There will now be $2\times N_p \times N_v$ points $\vec{v}$
($N_v$ such points for each of the $2\times N_p$ extremal boundary points in
$\{\hat{B}_\text{MaxMin}(\vec{M})\}$).
\\
\item(5B) At each point $\vec{v}$ above, find the $N_g$ nearest neighbor points as in step (2A) above
and evaluate equation (\ref{eqn:sstat}).  
Find the point $\vec{v}$ which maximizes $S$ and evaluate $\vec{\theta}\rightarrow\chi^2$ at
that point.  If
$\chi^2 \neq \chi^2_\text{lim}$, use bisection to find a new boundary point.
Store the resulting $\chi^2=\chi^2_\text{lim}$ point in $\{\hat{B}(\vec{M})\}$
\end{itemize}

The modification suggested in section \ref{sec:focus} is designed to explore 
$\chi^2=\chi^2_\text{lim}$ surface by starting from its extreme corners and stepping
perpendicularly to its ``radii'' (the vector $\vec{b}-\vec{M}$ would be a radius of
$\chi^2=\chi^2_\text{lim}$ described a sphere), with the hope of being able
to extend the bounds of $\{\hat{B}_\text{MaxMin}(\vec{M})\}$.  In this way,
our extended APS attempts to plum the full Frequentist confidence limit $\chi^2\le\chi^2_\text{lim}$
as rapidly as possible.  However, we do not want this to come at the expense of the wide-ranging
exploration of steps (1A-3A).  Our version of APS alternates between performing steps (1A-3A) and steps
(1B-5B).  If there are $N_m$ distinct known local minima in $\{\hat{\Gamma}\}$, 
our algoritm will perform
$N_m$ iterations of steps (1A-3A), and then perform steps (1B-5B) once for each
$\vec{M}$ in $\{\hat{\Gamma}\}$.

@@

The penultimate modification to the original APS algorithm which we propose in this
work is to utilize function minimization to identify local minima in $\chi^2$.
One of the principal drawbacks of the algorithm presented in steps (1A-3A) is
that it makes no account of ``near misses.''  If step (3A) finds that $\chi^2$
is lower than average but still greater than $\chi^2_\text{lim}$, the original
APS algorithm does nothing with that information.  We have tried to improve upon
that with our incorporation of bisection in sections \ref{sec:focus} and
\ref{sec:unitSphere} above.  An additional innovation is to allow APS to use points
with 
\begin{eqnarray}
\chi^2_\text{lim}&\le&\chi^2\le\chi^2_\text{threshold}\label{eqn:simplexCriterion}\\
\chi_\text{threshold}&\equiv&\chi^2_\text{min}+G\times(\chi^2_{1/10}-\chi^2_\text{min})
\label{eqn:threshold}
\end{eqnarray}
to seed a simplex function minimizer as developed by Nelder and Mead (1965).
Here, $G$ is a user-set parameter, $\chi^2_\text{min}$ is the minimum discovered
$\chi^2$ value and $\chi^2_{1/10}$ is the 1/10 quantile of the $\chi^2$ values
discovered by steps (1A-3A) (not those discovered by any of the modifications
discussed in sections \ref{sec:focus}, \ref{sec:unitSphere} or
\ref{sec:simplex}).

@@

\subsubsection{Simplex-driven APS search}
\label{sec:simplex_strad}

{\bf %marking new text

Having introduced the concept of the Nelder-Mead (1965) simplex search in
Section \ref{sec:simplex} above, we additionally use this tool to improve upon
the original APS search algorith, steps (1A)-(3A).  Broadly speaking, the
purpose of the original APS algorithm is to identify the point which maximizes
the $S$ statistic in equation (\ref{eqn:sstat}) and evaluate
$\vec{\theta}\rightarrow\chi^2$ at that point.  As more points are sampled, $S$
will, presumably, come to be maximized at points that are near the
$\chi^2=\chi^2_\text{lim}$ contour, but far from other sampled points, so APS
will efficiently explore the parameter space.  Steps (1A)-(3A) approximately
achieve this end by proposing a random sample of candidate points and evaluating
$\vec{\theta}\rightarrow\chi^2$ at the point which maximizes $S$ within that
sample.  We use the Nelder-Mead (1965) simplex to replace this process as follows:
\begin{itemize}
\item(1\~A) Perform step (1A) as above.
\\
\item(2\~A) Generate $N_p+1$ randomly chosen points $\{\hat{R}\}$ in parameter space.  
For each point $\vec{r}$ in $\{\hat{R}\}$, use equation (\ref{eqn:sstat}) as in steps 
(2A-3A) above to calculate $S$.  This set of $S$ values will be called
$\{S\}$.
\\
\item(3\~A) Use $\{\hat{R}\}$ and $\{S\}$ as the seed for a
Nelder-Mead (1965) simplex search.  At each subsequent point sampled by this
search, calculate $S$ as above.  Run this simplex until $S$ converges to a local maximum.
\\
\item(4\~A) Evaluate $\vec{\theta}\rightarrow\chi^2$ at the point which
maximized $S$.  Add this point to the list of points sampled by APS.  Return to
step (2\~A).  Repeat until convergence.
\end{itemize}
By choosing the maximum $S$ point based on a simplex search rather than a
finite random sample, we hope to find a true local maximum in $S$, rather than
just the maximum relative to the limited number of sample points chosen.  Steps
(1\~A)-(4\~A) replace steps (1A)-(3A) in our expanded APS algorithm.

} %marking new text

@@

We make code implementing our extended APS for 
an arbitrary $\vec{\theta}\rightarrow\chi^2$
function available at \verb|https://github.com/uwssg/APS/|.
The code is written in C++.  It is designed for users to use
with any $\vec{\theta}\rightarrow\chi^2$ function.  Users need only
to write a sub-class of the \verb|chisquared| class defined in
\verb|chisq.h| and \verb|chisq.cpp| which evaluates their desired
$\vec{\theta}\rightarrow\chi^2$ and pass an instantiation of this
sub-class to an instantiation of the \verb|aps| class defined in
\verb|aps.h| and \verb|aps.cpp|.  It is this \verb|aps| class which
does the work of performing the searches described above.
The user should not need to modify this class (though she is 
certainly welcome to).

Our \verb|aps| class provides the option for several
user-specified parameters.  We list them below along with the C++ 
methods (belonging to the class \verb|aps|) used to set them.

\begin{itemize}
\item The parameter $G$ from equation (\ref{eqn:threshold}) 
above can be set using the method
\verb|aps.set_grat|
\\
\item $\chi^2_\text{lim}$ can be manually set using the 
method \verb|aps.set_target|
\\
\item $\Delta \chi^2$ can be set in the constructor for the class \verb|aps|.
\\
\item The number $N_g$ of nearest neighbors used in 
steps (2\~A)-(3\~A) can be set in the constructor
for the class \verb|aps|
\\
\item The minimum and maximum bounds in parameter space can be set using the 
method \verb|aps.set_max| and \verb|aps.set_min|.  These bounds will constrain the range of
points sampled in step (2\~A)-(3\~A).  
They will not contrain searches performed with bisection or the 
simplex minimizer.
\\
\item If there is a characteristic length scale associated with a parameter, this can be set
using the method \verb|aps.set_characteristic_length|.  This length scale will be used to
normalize all distances in parameter space.  If it is not set, it will default to the
difference between the maximum and minimum values set by \verb|aps.set_max| and
\verb|aps.set_min|.  Setting a characteristic length scale will principally
effect what points are considered nearest neighbors for the purposes of steps
(2\~A)-(3\~A).
If the characteristic length scale of a parameter is large, that parameter will
be down-weighted when calculating the parameter-space distance between two points.
If the characteristic length scale of a parameter is small, that parameter will
be up-weighted in length calculations.
\end{itemize}

@@

\item(2D) Select the $3N_p+1$ nearest neighbor points to $\vec{a}$ from
$\{\hat{A}\}$ (this will obviously include $\vec{a}$ itself, which can be discarded).  Call
this set of nearest neighbors $\{\hat{N}\}$.
\\
\item(3D) For each point $\vec{n}$ in $\{\hat{N}\}$, calculate the vector
$\vec{\delta}$ such that $\delta_i=|n_i-a_i|$ (here the vertical bars denote the absolute value).  
For each component
$\delta_i$ of $\vec{\delta}$, starting with the largest and working towards the smallest:
\\
\begin{itemize}
\item(3Da) If $n_i>a_i$ and $a_{\text{max},i}$ has not been set, set $a_{\text{max},i}=n_i$.
\\
\item(3Db) If $n_i<a_i$ and $a_{\text{min},i}$ has not been set, set $a_{\text{min},i}=n_i$.
\\
\item(3Dc) If both $a_\text{max,i}$ and $a_\text{min,i}$ have already been set, move on to the
next largest $\delta_i$ and repeat at step (3Da).
\end{itemize}
\item(4D) Choose the next neighbor $\vec{n}$ and repeat from step (3D). Continue until all
$a_{\text{max},i}$ and $a_{\text{min},i}$ have been set.  Once all $a_{\text{max},i}$ and
$a_{\text{min},i}$ have been set, move on to step (5D).
\\
\item(5D) $\vec{a}_\text{max}$ and $\vec{a}_\text{min}$ should now describe an asymmetric,
rectangular hyperbox about $\vec{a}$.  To symmetrize it, step through all $N_p$ dimensions
and reset either $a_{\text{max},i}$ or $a_{\text{min},i}$ so that 
$d_\text{min}=a_i-a_{\text{min},i}$ and
$d_\text{max}=a_{\text{max},i}-a_i$ are equal to the minimum of $\{d_\text{max},d_\text{min}\}$
from the asymmetric hyperbox.  Return to step (1D) and repeat for the next $\vec{a}$.
\end{itemize}

At this point, each $\vec{a}$ in $\{\hat{A}\}$ should have a corresponding symmetric hyperbox
surrounding it described by the corresponding 
$\vec{a}_\text{max}(\vec{a})$ and $\vec{a}_\text{min}(\vec{a})$.  
To integrate the posterior probability density, we will assume that all of the
parameter space points in each hyperbox correspond have $\chi^2=\chi^2(\vec{a})$,
{\bf %marking new text
i.e. the $\chi^2$ discovered by APS for the point at the center of the hyperbox.
} %marking new text
The posterior probability associated with each box is then
\begin{eqnarray}
P(\vec{a})&=&\frac{V(\vec{a})}{P_\text{tot}}\times\exp\left[-\chi^2(\vec{a})/2\right]
\label{eqn:posterior}\\
P_\text{tot}&\equiv&\sum_{\vec{a}} V(\vec{a})\times\exp\left[-\chi^2(\vec{a})/2\right]
\end{eqnarray}
where $V(\vec{a})$ is the hypervolume of the hyperbox surrounding $\vec{a}$.
To plot the $(1-\alpha)\%$ credible limit, one need only plot the points in hyperspace that
enclose the lowest-$\chi^2$ hyperboxes containing $(1-\alpha)\%$ of the total probability.
We test this algorithm on a function with an exactly known credible limit
in Section \ref{sec:toy} (see Figures \ref{fig:toyBayes2}-\ref{fig:toyBayes4}) 
and on a real, physical dataset in Section \ref{sec:wmap} (see Figure \ref{fig:contours_bayes}).  
In  Section \ref{sec:toy}, we find that our results agree with the known credible limit.
{\bf %marking new text
In Section \ref{sec:wmap} we find that APS correctly identifies the region and general
shape of the Bayesian credible limits, however, with significant noise.  The Bayesian
credible limits found by APS are systematically larger than the Bayesian credible
limits found by MCMC.  This motivates our claim in the introduction that APS may be
principally useful for identifying regions of high posterior density on largely unknown
likelihood functions, allowing researchers to more shrewdly set thep priors for sampling
algorithms like MCMC.
}%marking nex text

@@

Steps (2\~A)-(3\~A) of \APS requires that a prediction $\mu$ be made regarding the value
of $\chi^2$ at points as yet unsampled by the algorithm.  An uncertainty $\sigma$
must also be assigned to this prediction.  APS writ large is agnostic regarding
the method used to find $\mu$ and $\sigma$.  We choose to use a Gaussian process.

@@

The normalization constant $K$ in equation (\ref{eqn:likelihood}) -- known as the
``Kriging parameter'' for the geophysicist who pioneered the overall method -- 
also
must be assumed.  Determining the value of $K$ 
is somewhat problematic because, examining equation
(\ref{mu}), one sees that the factors of $K$ and $K^{-1}$
cancel out of
the prediction $\mu$, so that the assumed value of $K$ has no effect on the accuracy
of the prediction.  
If the opposite had been true, one could heuristically set $K$ to
maximize the accuracy of $\mu$.  
Instead, we set $K$ to the value that maximizes the likelihood
of the input data, i.e. the $N_g$ nearest neighbors in parameter space
used to perform the Gaussian Process inference.

@@

As APS samples more points in parameter space, we expect it to learn more about the
$\vec{\theta}\rightarrow\chi^2$ function and thus be capable of a more accurate Gaussian
Process, so we allow it to periodically reset the hyper-parameters $\vec{h}$.  However,
minimizing $E$ can be a very time-consuming process, so we have APS keep track of the amount of
time elapsed since the algorithm began.  We divide this value by the number of calls made to
the $\vec{\theta}\rightarrow\chi^2$ function and 
{\bf %marking new text
compare this 
average time per $\vec{\theta}\rightarrow\chi^2$ evaluation with all of the overhead
imposed by the APS algorithm to the average
amount of time spent on one call to $\vec{\theta}\rightarrow\chi^2$ without the overhead
imposed by APS. 
}%marking new text
APS us only allowed to
reset its hyper-parameters if the difference between these two times is less than 0.1 
second per call to $\vec{\theta}\rightarrow\chi^2$
(of course, we do require that APS optimize its hyper parameters at least once, so that
we can have confidence in our Gaussian Process model).
